Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2018-03-06 article3d.Rmd
clean figures

The estimated medians are

```{r estmed}
knitr::kable(round(medi, 1))
```

@@

```{r forty64}
source("R/getWeights.R")
da <- vb1[vb1$sex == "male" & vb1$period == "2005-2013", ]
da$wght <- getWeights(da)
##
fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), data = da)
med <- summary(survfit(fit))$table[, "median"]
fitw <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = da)
medw <- summary(survfit(fitw))$table[, "median"]
oldpar <- par(mfrow = c(1, 2))
plot(survfit(fit), mark.time = FALSE, col = scol, lty = 1:5, xlim = c(40, 65), 
     xlab = "Age", ylab = "Survival", main = "Unweighted", axes = FALSE)
tics <- c(40, 45, 50, 55, 60, 65)
axis(1, at = tics)
axis(2, at = c(0, 0.5, 1), las = 1)
box()
abline(h = 0)
abline(h = 0.5, lty = 3)
legend("bottomleft", legend = levels(vb2$hisclass), col = scol, lty = 1:5)
##lines(c(med["none"], med["none"]), c(0, 0.5), col = "red", lty = 3)
##lines(c(med["elite"], med["elite"]), c(0, 0.5), col = "black", lty = 3)
plot(survfit(fitw), mark.time = FALSE, col = scol, lty = 1:5, xlim = c(40, 65), 
     xlab = "Age", ylab = "Survival", main = "Weighted", axes = FALSE)
tics <- c(40, 45, 50, 55, 60, 65)
axis(1, at = tics)
axis(2, at = c(0, 0.5, 1), las = 1)
box()
abline(h = 0)
abline(h = 0.5, lty = 3)
legend("bottomleft", legend = levels(vb2$hisclass), col = scol, lty = 1:5)
```

@@

# The effect of social class on mortality

@@

```{r cumsw, fig.cap = cap0, fig.scap = scap0, fig.height = 8, out.height = "14cm"}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
midper <- c(1812.5, 1837.5, 1862.5, 1887.5, 1912.5, 1937.5, 1981, 1991, 2000.5, 2009.5)
plot(midper, resw[1, ], col = scol[1], type = "b", ylim = c(58, 65), xlab = "Period", 
     ylab = "Age", axes = FALSE, main = "Women, 40-64")

axis(1, at = midper, labels = labb)
axis(2, las = 1)
box()
for (i in 2:5){
    lines(midper, resw[i, ], col = scol[i], type = "b")
}
plot(midper, resm[1, ], col = scol[1], type = "b", ylim = c(58, 65), xlab = "Period", 
     ylab = "Age", axes = FALSE, main = "Men, 40-64")
axis(1, at = midper, labels = labb)
axis(2, las = 1)
box()
for (i in 2:5){
    lines(midper, resm[i, ], col = scol[i], type = "b")
}
##woutp <- plotED.glm(resw, col = scol, logScale = TRUE, ylab = "H(65|30)",
  ##              main = "Women 30-64", ylim = c(0, 1))##, exclude = 5:6)
##moutp <- plotED.glm(resm, col = scol, logScale = TRUE, ylab = "H(65|30)",
  ##              main = "Men 30-64", ylim = c(0, 1))##, exclude = 5:6)
par(oldpar)
```

>>>>

The estimated medians are (see Table \@ref(tab:estmed).

```{r estmed}
knitr::kable(round(medi, 1), caption = "Estimated medians, men 65-99, 2005-2013.")
```

@@

```{r forty64, fig.cap = "Survival curves, men 40-65, 2005-2013.", fig.height = 4}
source("R/getWeights.R")
da <- vb1[vb1$sex == "male" & vb1$period == "2005-2013", ]
da$wght <- getWeights(da)
##
fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), data = da)
med <- summary(survfit(fit))$table[, "median"]
fitw <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = da)
medw <- summary(survfit(fitw))$table[, "median"]
oldpar <- par(mfrow = c(1, 2))
plot(survfit(fit), mark.time = FALSE, col = scol, lty = 1:5, xlim = c(40, 65), 
     xlab = "Age", ylab = "Survival", main = "Unweighted", axes = FALSE)
tics <- c(40, 45, 50, 55, 60, 65)
axis(1, at = tics)
axis(2, at = c(0, 0.5, 1), las = 1)
box()
abline(h = 0)
abline(h = 0.5, lty = 3)
legend("bottomleft", legend = levels(vb2$hisclass), col = scol, lty = 1:5)
##lines(c(med["none"], med["none"]), c(0, 0.5), col = "red", lty = 3)
##lines(c(med["elite"], med["elite"]), c(0, 0.5), col = "black", lty = 3)
plot(survfit(fitw), mark.time = FALSE, col = scol, lty = 1:5, xlim = c(40, 65), 
     xlab = "Age", ylab = "Survival", main = "Weighted", axes = FALSE)
tics <- c(40, 45, 50, 55, 60, 65)
axis(1, at = tics)
axis(2, at = c(0, 0.5, 1), las = 1)
box()
abline(h = 0)
abline(h = 0.5, lty = 3)
legend("bottomleft", legend = levels(vb2$hisclass), col = scol, lty = 1:5)
```

@@

@@

```{r cumsw, fig.cap = "Expected years of survival until the age of 65, given survival to age 40, by period and hisclass.", fig.scap = scap0, fig.height = 8, out.height = "14cm"}
source("R/plotED.glm.R")
oldpar <- par(mfrow = c(2, 1))
midper <- c(1812.5, 1837.5, 1862.5, 1887.5, 1912.5, 1937.5, 1981, 1991, 2000.5, 2009.5)
plot(midper, resw[1, ], col = scol[1], type = "b", ylim = c(58, 65), xlab = "Period", 
     ylab = "Years", axes = FALSE, main = "Women, 40-64", lty = 1, pch = 1)

axis(1, at = midper, labels = labb, las = 2, cex.axis = 0.5)
axis(2, las = 1)
abline(h = 65)
box()
for (i in 2:5){
    lines(midper, resw[i, ], col = scol[i], type = "b", pch = i, lty = i)
}
legend("bottomright", legend = his, col = scol, lty = 1:5, pch = 1:5, cex = 0.7)
plot(midper, resm[1, ], col = scol[1], type = "b", ylim = c(58, 65), xlab = "Period", 
     ylab = "Years", axes = FALSE, main = "Men, 40-64", lty = 1, pch = 1)
axis(1, at = midper, labels = labb, las = 2, cex.axis = 0.5)
axis(2, las = 1)
abline(h = 65)
box()
for (i in 2:5){
    lines(midper, resm[i, ], col = scol[i], type = "b", lty = i, pch = i)
}
legend("bottomright", legend = his, col = scol, lty = 1:5, pch = 1:5, cex = 0.7)
##woutp <- plotED.glm(resw, col = scol, logScale = TRUE, ylab = "H(65|30)",
  ##              main = "Women 30-64", ylim = c(0, 1))##, exclude = 5:6)
##moutp <- plotED.glm(resm, col = scol, logScale = TRUE, ylab = "H(65|30)",
  ##              main = "Men 30-64", ylim = c(0, 1))##, exclude = 5:6)
par(oldpar)
```
